---
title: ""

knit: (function(input_file, encoding) {
  out_dir <- 'docs';
  rmarkdown::render(input_file,
 encoding=encoding,
 output_file=file.path(dirname(input_file), out_dir, 'example3.html'))})
---

## Example 3 Adjusting subcatchment paramters after installing GIs

### 1. Overview

In this example, the function to modify subcatchment paramters after modeling GIs externally is explored. Additional changes to the catchment can be passed to the `modify_inp` function through the *GI_plan* variable, where tibble columns with names in the format *section_variable* are defined. The *"section"* term corresponds to the name of the tab used in swmm input files, such as the *INFILTRATION* tab, the the infiltration parameters are store under this tab. The *"variable"* term corresponds to the name of the variable, e.g., *"MaxRate"* corresponds to the maximum infiltration rate. Thus, the column can be named as *"INFILTRATION_MaxRate"*, and different values can be assigned to the column to reflect the changes to subcatchment parameters after installing the GI associated with the inflows specified in the *inflow_path* column (see example 1 for more explanations).

The goal of this example is to show the consequences of changing subcatchment width after modeling bioretention cells and their controlled areas externally.

The catchment studied in this example is the same as in example 2. It has three subcatchments, as shown in the figure below.

```{r warning=FALSE, message=FALSE, results="hide"}

# Load toolbox functions
source("interface_functions.R")

inp <- read_inp("./example/example3/raw_catchment.inp")

# Functions provided in swmmr are used for visulization
sub_sf <- subcatchments_to_sf(inp)
lin_sf <- links_to_sf(inp)
jun_sf <- junctions_to_sf(inp)
rg_sf <- raingages_to_sf(inp)

lab_coord <- sub_sf %>% 
  sf::st_centroid() %>%
  sf::st_coordinates() %>% 
  tibble::as_tibble()

lab_rg_coord <- rg_sf %>% 
  {sf::st_coordinates(.) + 500} %>% # add offset
  tibble::as_tibble()

sub_sf <- dplyr::bind_cols(sub_sf, lab_coord)
rg_sf <- dplyr::bind_cols(rg_sf, lab_rg_coord)

ggplot() + 
  # first plot the subcatchment and colour continously by Area
  geom_sf(data = sub_sf, aes(fill = Perc_Imperv), color = "black", alpha = 0.8) + 
  scale_fill_gradient(low = "lightgreen", high = "indianred4") +
  geom_sf(data = lin_sf, colour = "blue", size = 1) +
  geom_sf(data = jun_sf, colour = "grey20", size = 2) +
  geom_label(data = sub_sf, aes(X, Y, label = Name), size = 3) +
  labs(x = "x-axis",
       y = "y-axis",
       fill = "Imperviousness\n(%)") +
  scale_fill_gradient(limits = c(0,100), low = "#009E73", high = "#D16103") +
  theme_bw(base_size = 10)
```

### 2. Evaluate the consequences of changing subcatchment width and Manning's n after modeling bioretention cells exterally.

Three units of bioretention cells are assumed to be implemented in each subcatchments. The design option 2 is used, which corresponds to the outflow hydrograph stored in file *bc2_outflow.txt*. More details on bioretention cell designs can be found in example 2.

The following code chunk creates multiple tibbles stored in a list. Each tibble correspond to a bioretention cell modeling scenario, and for each scenario the subcatchment width of the subcatchments is set to a random value between 30 and 500, and Manning's n of the impervious areas of the subcatchment is set to a random value between 0.01 and 0.02. Manning's n is a measure of surface roughness. Multiple random scenarios are evaluated to assess the consequences of changing these parameters. The information on catchment width is provide in the *width_adj* column, and on Manning's n is provided in the *SUBAREAS_N-Imperv* column.

```{r}
set.seed(10)

inflow_path <- paste0("./example/example3/bc", 2, "_outflow.txt")

N = 50 # Number of scenarios
gi_plans <- vector("list", 10) # list storing the random GI install plan
for (i in 1:N){
  gi_plans[[i]] <- tibble(
    inflow_path = rep(inflow_path, 9),
    outlet = rep(0, 9),
    subcatchment_name = rep(c("S1", "S2", "S3"), each = 3),
    per_area_rep = rep(0, 9),
    imp_area_rep = rep(0.1, 9),
    width_adj = rep(runif(3, min = 30, max = 500), each = 3),
    `SUBAREAS_N-Imperv` = rep(runif(3, min = 0.01, max = 0.02), each = 3)
  )
}
```

The scenarios are evlauted usuing the `write_routing_interface_file` and `modify_inp` functions provided by the toolbox and the swmmr functions.

```{r warning=FALSE, message=FALSE, results="hide"}
read_outflow <- function(fpath = "outflow.txt"){
  # Purpose: read simulated outflow hydrograph
  # Input: 
  #    fpath = file path of the simulated outflow, which is defined in the [FILES] tabe in SWMM input file
  # Output:
  #    a tibble stores the outflow hydrograph
  
  read_table(fpath, skip = 7) %>%
    transmute(datetime = ymd_hms(paste(Year, Mon, Day, Hr, Min, Sec)),
            flow = FLOW) %>%
    arrange(datetime)
}

# Read the SWMM input file for the original subcatchment
inp <- read_inp("./example/example3/raw_catchment.inp")

# Create a list to store the simulated outflow hydrographs
outflows <- vector("list", N)
for (i in 1:N){
  
  # create a routing interface file
  gi_plan <- gi_plans[[i]]
  routing_interface_path <- "./example/example3/routing_interface.txt"
  write_routing_interface_file(
    GI_plan = gi_plan, 
    inp = inp, 
    routing_interface_path = routing_interface_path
  )
  
  # modify the object associated with swmm input file
  new_inp <- modify_inp(
    GI_plan = gi_plan, 
    inp = inp, 
    routing_interface_path = routing_interface_path
  )
  
  # write the new SWMM input file
  new_inp_path <- "./example/example3/test.inp"
  write_inp(new_inp, file = new_inp_path)
  
  # run simulation
  run_swmm(new_inp_path)
  
  # get simulated outflow hydrograph
  outflows[[i]] <- read_outflow() %>%
    mutate(case = i)
}

# join the simulated hydrograph into a single tibble
outflows <- outflows %>%
  bind_rows()
```


### 3. Assessment results

The following code generates a figure comparing the hydrographs generated in each scenario. The grey lines correspond to the results for each scenario, and the blue line shows the mean values. Considerable varations between different hydrographs can be observed, especially in the predicted peak flows.

```{r}
# Keep data for relatively high flow periods
data_plot <- outflows %>%
  filter(datetime <= ymd_hm("2020-06-02 00:00"))

# Get the mean hydrograph of all the scenarios
data_plot2 <- outflows %>%
  filter(datetime <= ymd_hm("2020-06-02 00:00")) %>%
  group_by(datetime) %>%
  mutate(mean_flow = mean(flow)) %>%
  ungroup()

# Ploting
ggplot() +
  geom_line(data = data_plot, aes(datetime, flow, group = case), color = "grey50", size = 0.5) +
  geom_line(data = data_plot2, aes(datetime, mean_flow), color = "blue", size = 0.8) +
  labs(x = "Time",
       y = "Flow rate [CFS]") +
  theme_bw()
```


